---
title: Exploratory Data Analysis
format: typst
---

### Question 1

Which  district has the most unequal distribution of precinct sizes?

### Answer 1
```{r}
nice |>
  group_by(CDDIST) |>
  summarise(sd_turnout = sd(TOTVOTE, na.rm = TRUE)) |>
  arrange(desc(sd_turnout))

nice |> filter(CDDIST != 0)
```

I used the cleaned nice dataset and looked at how much precinct turnout varied within each congressional district by calculating the standard deviation of TOTVOTE. Then I saw that District 36 stood out with the highest variation in precinct size (about 1,061) while District 12 had a similarly wide spread (around 1,001), finally followed by District 32 (roughly 895).

### Question 2

Which  one of the precincts had the most votes for a 3rd party in 2024?

### Answer 2

```{r}
nice |>
  mutate(tp = PRSLIB01 + PRSAIP01 + PRSGRN01,
         s = tp / TOTVOTE) |>
  slice_max(s, n = 1)
```

 I calculated the share of presidential votes going to third-party candidates and found that the highest share came from a very small precinct in Tulare County (FIPS 06107), which was only 147 votes anyway.

### Question 3

Which one od the precincts had the closest  race in absolute vote margin?

### Answer 3

```{r}
nice |>
  filter(PRSDEM01 > 0, PRSREP01 > 0) |>
  mutate(margin = abs(PRSDEM01 - PRSREP01),
         total_pres = PRSDEM01 + PRSREP01) |>
  filter(margin > 0) |>
  slice_min(margin, with_ties = FALSE)
```

Put your answer here.